### 2021-07-20 2.17-plagiarise.Rmd

# Making sure that you don’t plagiarise {#plagiarise}

Plagiarism is when you copy somebody else's work. We're most familiar with plagiarism of writing and these days this is especially easy with the copy and paste function. Most people are not even aware that they have plagiarised because sometime in the past they copied and pasted the work of somebody else into a document and this later became incorporated into their text without them being aware of it.

Plagiarism is a problem because, essentially, you are taking somebody else's work without attributing it to them.

## How to know if you have plagiarised

Today there are several pieces of software that are used to scan text that's written and available on the internet to discover plagiarism. One such example is TurnItIn. One of the outputs of TurnItIn is to highlight text that matches other text already on the internet. As almost all journals publish on the internet, TurnItIn can accurately determine if text has been copied from another article or website. I usually set TurnItIn to determine plagiarism with five or more consecutive words.

Remarkably it is very difficult to come up with exactly the same words that someone else used to describe a phenomenon. Most people when they think about it feel that it wouldn't be that surprising if they came up with exactly the same words as somebody else.

When caught plagiarising most undergraduates claim that they simply read an article and then later happen to write the same words that were in the article. They categorically deny that they ever copied or pasted text from the article into their work.

__Try it. When you try it you will learn what plagiarism is all about.__

Read an article, and then try to write text that is exactly the same as that in the article without looking back at the article itself.

Unless you have an eidetic memory you will fail at this task.

(ref:plagiarism) **It's always tempting to copy and paste, but it is likely to lead to plagiarism.** Copy (Ctrl + c) and paste (Ctrl + v) have become so easy that it is tempting to pick up portions of appropriate text directly from papers and then slot them into our own work (Figure \@ref(fig:plagiarism)). However, this is plagiarism and can easily be found by using software like TurnItIn. Most institutions will require checks for plagiarism on your thesis after submission, with dire consequences if your text fails.

```{r plagiarism, echo=FALSE, out.width = '95%',  fig.cap="(ref:plagiarism)"}
knitr::include_graphics('figures/plagiarism.png')
```

This is not to say that no 5 or more words can ever be the same as someone else's. There are situations in which this happens. Think of addresses, quotes, certain laboratory equipment or protocols, and certainly references at the end of your paper. So there are many times when TurnItIn will come back with matching text. This is not what we're looking for in plagiarism.

##  What to do if plagiarism is detected in your work

It's remarkably easy to remove plagiarism from your work.

__Here's what you do:__

- Read the sentence that has been plagiarised several times to yourself. 
- Now without looking at that sentence, write another sentence that has the same meaning. Because it's very hard to replicate somebody else's words without copying them, what you should find is that you've written a fresh unplagiarised sentence. 
- This can now be added to your text, changed as appropriate to fit your existing text. And that should be the end of your plagiarism worries.

__Here's what not to do:__

- Take the sentence swap out some of the words for synonyms and pass it off as your own. 
- The sentence will still have the same structure that you copied and, essentially, this is still somebody else's work. Moreover, TurnItIn will still recognise this as plagiarism.

## How can you make sure that you never plagiarise?
Quite simply, if you never cut and paste, you won’t plagiarise. It’s that easy. I understand why people copy and paste as a way to get started, or because someone else has written something so well, it’s hard to believe that you could ever write it any better. But actually, you can write it just as well, and writing it in your own words is worth so much more.

Don’t forget that the penalty for plagiarism in your thesis might well be that you fail.

It still seems amazing to me when I submit a student manuscript to TurnItIn and see that it is completely free of any plagiarism. There are so many words in English, and so many different ways of putting them that you really can have your own writing style. Your writing style will be as unique to you as a fingerprint, and it will be entirely free of plagiarism. It's something you can celebrate.

Want to read more about plagiarism to understand what it is: see Louw [-@louw2017defining]

## What is autoplagiarism?

I would maintain that there is not really any grey area in plagiarism. So what are people talking about when they refer to a grey zone? The nominal grey zone within plagiarism comes when you copy your own text, this is sometimes referred to as 'text-recycling', 'self-plagiarism' or 'autoplagiarism'. A new guideline from Hall et al. [-@hall2021understanding] sets out the different types of grey plagiarism, or text recycling:

### Developmental recycling {#develop}

Developmental recycling is when you are reusing text that you have written for example between your proposal and something you intend for publication or in an ethics application that you also want to use in your thesis. All of this sort of developmental recycling is permitted and actually encouraged. I would further encourage you to use the opportunity of recycling this text to develop it and refine it further, condensing and improving where you can.

### Generative recycling {#generative}

Generative recycling is where you take pieces of already published text for example from the methods when it does not make sense to change the text or actually makes it more obscure to reword it in order to avoid plagiarism. In my experience this doesn't amount to more than a few sentences describing technical settings on equipment. However this will depend strongly on your own subject area and may amount to larger chunks of text. I suggest that it is usually possible to reword most of the methods sections of papers. You should really only be generatively recycling material if you cannot avoid it: i.e. when the text becomes more obscure by your attempts to reword it.

### Adaptive recycling

Adaptive recycling is where you are using published text as the basis for a different type of content (e.g. a popular article online, a magazine, or op-ed). I think that this kind of text recycling is quite unnecessary because you almost certainly need to reword your text for a different audience. There maybe times such as figure legends where you need to reuse text that was already published. If you do find yourself in such a position then check with the copyright owner of the material that you are able to reuse the text that you want without legal issues.

### Duplicate recycling

Duplicate recycling (also known as 'manuscript recycling') is where large tracts of texts are essentially the same for the same message and audience. This is never likely to be sanctioned as it suggests that you are attempting to publish the same work twice. It will not be legal or ethical. Manuscript recycling is a bigger problem in other subject areas, notably economics [@geraldi2021self-plagiarism], publishing issues in economics appear to need a lot more professionalism.

## Overall - just don’t plagiarise

The simple way around this problem is not to plagiarise any of your own text (self-plagiarism) or any of anybody else's. This is, as I've explained above, relatively easy because as long as you don't copy and paste any text you'll find it very hard to actually use somebody else's words. However it can get a little bit tricky when you are writing the materials and methods section of a paper, and especially when that paper uses exactly the same materials and methods as a previous paper. The temptation is great to go and copy the text that you've written before. But be aware that, however grey, this is plagiarism and you should not do it.

The likelihood is that because you're using the same names and their relationship is staying fixed as shown in the example above in materials and methods you are more likely to come up with text that a piece of software like turn it in will show is plagiarised even when it isn't.