### 2021-07-20 4.00a-finishing.Rmd

# Now that you have finished your data chapters

In this last part of the book, I am going to presume that you have done all of the work for the data chapters, and written them up as distinct chapters of your thesis. When you look at PhDs that are stored in your institution's libraries, or those that you’ve referred to in the course of your study, you will notice that they have two additional chapters in addition to data chapters. These are often quite short (although lengths vary), and serve to introduce and conclude the PhD thesis (Figure \@ref(fig:thesis-construction). While these are not usually considered to be the ‘meat’ of the PhD itself, they are important sections to reflect on and carefully consider, as they attempt to pull together data chapters that are written as stand-alone documents into a [cohesive thesis](#criteria). To this end, they require the same careful planning that any other section in your thesis deserves, with one important condition: that they should not impede the timely submission of your thesis.

A typical construction of the data chapters in your thesis together with a introduction and conclusion. The exact construction of your thesis may vary, especially with the number of data chapters that you have. The relationship between thesis chapters may also not be linear as presented here.

(ref:thesis-construction) **A typical construction of the data chapters in your thesis together with a introduction and conclusion.** The exact construction of your thesis may vary, especially with the number of data chapters that you have. The relationship between thesis chapters may also not be linear as presented here.

```{r thesis-construction, echo=FALSE, fig.show='hold', fig.align='center', out.width='90%', fig.cap="(ref:thesis-construction)"}
knitr::include_graphics('figures/thesis-construction.png')
```

At this point, we have to acknowledge that different institutions require different deadlines of their students in order to finish their PhD studies and submit their theses for examination. In most institutions (to my knowledge) the time to submission is the most important, and can vary between 3 and 5 years (most institutions stipulate a minimum period of 2 years and some have no maximum duration). This period is critical with respect to funding. If funding ceases and your institution allows you to continue your studies, you may have to raise the funding to pay fees as well as to maintain your living costs. After submission, there will be a period of time while your thesis is in examination, and another period before the final graduation. Each of these milestones comes with its own set of deadlines and conditions. It is well worth being aware of all these issues well in advance of any submission deadline (ask your advisor). Important aspects for you to note are whether or not you need to remain registered at your institution after you have submitted your thesis, and whether or not your institution will require payment for your registration.

Whatever your deadlines are will dictate how much time (and energy) you have to devote to the opening and closing chapters of your thesis. They can become publishable in their own right, or at the bare minimum, they can serve to allow you to submit a completed thesis. Here I will focus on the latter, minimum requirement (in my eyes), with the understanding that you should dedicate as much time as you have available to produce a quality product that you can feel proud of.

The next chapters deal with writing the [introductory chapter](#PhDintro), the [concluding chapter](#PhDconc), and then finally [formatting your thesis](#formatting) for submission. These final hurdles require some time and thought. Given that not all of you will have all of the time that you need, this is often the time when the demand on your writing skills reaches a peak. If you are in this position, I urge you to use your time wisely and to plan carefully what you will write.