### 2021-09-05 1.00-Part1.Rmd

\mainmatter

# (PART) Part I: Getting ready for submission {-}

# Submitting your work to a journal {- #part1}

Part 1 of this book is about the publication arena. I provide simple explanations of what scientific journals are for, and who the gatekeepers are. The publishing world is at a turning point, and before you start publishing you should be aware of the current trends and what is possible.