### 2021-09-05 1.00a-closedtoopen.Rmd

# The transition from closed to open {#closedtoopen}

There are a lot of problems in publishing in the biological sciences, but it doesn't have to be this way. The aim for this book is firstly to help you navigate the current 'closed' system (while acknowledging that parts are already open), and act as a catalyst toward a more open, transparent and equal system for the future not only of publishing science, but permeating throughout the culture of the scientific project. We have all the tools to make this transition now, and I think that this change will likely come within the time frame of the careers of you as an Early Career Researcher. But as you will see, this change needs to be driven.

While I'm going to pitch the transition from closed to open publishing as a simple process, and as a move from darkness into light, I acknowledge that it might best be defined as a [wicked problem](#wicked). I hope that a lot of these complexities will come out in the book, but acknowledge that a lot will be left unsaid. A large unanswered question is what happens to all the downstream impacts of losing a lucrative academic publishing business that employs tens of thousands of people. I like to think that many of these skilled people, who themselves are not recipients of the large fees acquired by the publishers, would be absorbed into the repurposed institutional libraries. No doubt, there will be casualties. But my belief is that the importance of the scientific project, and the [wicked problem](#wicked) we currently face in academic publishing, far outweighs the problems that we will see during the transition.

Throughout this book, I will make reference to the '[scientific project](http://howtowriteaphd.com/lifescientific.html)' as a broader philosophical stance in which our studies in biological sciences is simply a part [see @measey2021how]. This is a stronger institution without the culture of assumed knowledge and elitism that we see today. I consider movement away from the current model of publishing to be at the heart of this essential institutional transformation.

## Three fundamentals of publishing in Biological Sciences

There are three fundamental concepts that lie at the heart of understanding of current publishing models in the Biological Sciences (Figure \@ref(fig:publishing-triangle)). In themselves, none of them should be particularly influential as they don't relate to your study, how well the study was done, or what your results were. Nevertheless, these three aspects of publishing are key in your understanding of the nuances of publishing, and your understanding will likely make the difference between publishing your work being an obstacle that is occasionally insurmountable, and finding your way with a lot more ease through the process.

### Gatekeepers {#gatekeepers}

The journal editor is the kingpin, and sits at the centre of this triangle, and has the power, backed by their gatekeeping advisory board and associate editors, to continue the current model, or oversee the change. [Editors make decisions](#editors), not simply whether to accept or reject your manuscript, but also to implement policies that take their journals in one direction or another. In some models, they are given this power (usually democratically) by the [scholarly society](#society) that they represent, and oversight is granted by an editorial board who mediate in any dispute, but also in theory have influence over the editorial policy. Editors appoint associates that handle many of the manuscripts that are submitted, shuffling them between reviewers and authors until they feel that they are worthy of publication (or not). These associate editors are also responsible for implementing the policies of the editor, the editorial board, and the scholarly society. In models where there is only a (for-profit) publisher, the publisher appoints the editor and together they appoint the editorial board. In both models, the editorial board, editors and associate editors are thought of as being the gatekeepers to the scholarly integrity that permeates scientific publishing. There is more information on the advisory and editorial boards in [Part I](#boards).

(ref:publishing-triangle) **A simple schematic for the three fundamentals of publishing in the current Biological Sciences model**. Citations, peer review and impact factors each have direct impact on each other and your understanding of each one and how they relate to the other will be pivotal in clarifying your understanding of how to publish your work. At the heart of the process are the gatekeepers: Editorial board, associate editors and the editor.

```{r publishing-triangle, echo=FALSE, out.width = '85%', fig.cap="(ref:publishing-triangle)"}
 knitr::include_graphics("figures/publishing-triangle.png")
```

Gatekeeping takes a lot of time and effort, and there are plenty of places where the current system lacks the transparency that is needed. In theory, there's nothing wrong with this model, were it not for the dark forces of for-profit publishers that have used the system for their own gain. Instead of the gatekeepers focussing on science, they have become distracted by metrics and hype promoted by publishers. Each of the three facets that surround the gatekeepers in the current publishing model (Figure \@ref(fig:publishing-triangle)) need to be changed to open up the system for a more equitable future, and eliminate the current biases that favour the select few.

### Is it possible to purge the publishers?

Editors don't have a complete free reign over what to do with their journal. They may be constrained by the publishers, if their journal is run on a for-profit model, or by a contract with a publisher, for most society journals. Journals which are independent of societies and publishers are very rare, but do exist. However they work, editors sit as king-pins of the system. There are examples of editors who have taken all of their associate editors and authors and moved their entire platform to a not-for-profit system. This has also meant changing the name of the journal, as the publisher often owns this. The first example, that I'm aware of, happened in 2015 in the social sciences where the editor of _Lingua_ walked away from publishers Elsevier [see @bakovic2017language]. As Elsevier owned the name, the editor, Johan Rooryck, started a new Open Access journal [_Glossa_](https://www.glossa-journal.org/). There was a fight (see [Rooryck's website](https://www.rooryck.org/interaction-with-elsevier)), but Rooryck showed that it could be done, and has therefore paved the way for others. Not only did Rooryck show proof of concept, but he formed the [Fair Open Access Alliance](https://www.fairopenaccess.org/), who have managed to pull 6 titles away from Elsevier since _Lingua_ flipped. It is important to add here that when the gatekeepers leave, the publisher simply approaches new people to take over, and the old journals remain alongside the new. It appears that there is no shortage of academics who are prepared to step into the shoes of editors who part company with the publishers. See what draws editors in [here](#boards).

All editors are answerable to their advisory board, and to the scholarly society from which they were (usually) voted into office. The society (often through the editor) signs a deal with the publisher, and this usually runs for a period of 5 years (see [Part IV](#money)). Society journals usually own their own content and title, and so can decide to leave the publisher whenever their contract expires. It just takes will power, and being prepared to say goodbye to that [income stream](#money).

**So why do the gatekeepers stay with the publishers?**
It's mostly smoke and mirrors. Editors and gatekeepers in general are busy people. Their gatekeeping roles are not their primary jobs (for the most part), and if they are then the publishers are paying their wages. What they are most interested in is a smooth system that works with minimum effort on their part. This is what the publishers have established so well, and the principle way in which they will try to persuade gatekeepers to stay with them. Next is the [money](#money), which flows from the publishers into the accounts of the societies, with occasional small amounts to editors (and in rare cases associate editors) as expenses. Such perks used to be more substantial, like trips to conferences, hotel stays and wining and dining. But I understand that this is largely gone now. The perceived prestige and professional advancement are [discussed in detail later for all gatekeepers](#boards). Last is the inertia on the part of gatekeepers to change, as they don't experience the pain of the authors or the libraries that pour money into the publishers. This is a reason why bringing societies into closer contact with institutions is an important step in the process.

## Peer review
Few would argue that peer review is at the heart of publishing in science today. Consequently, there are several chapters in this book concerned with the subject: [What is peer review](#peerreview); [what to expect from your peer reviewers](#peerreview2); [how to respond to peer review with a rebuttal](#rebuttal); [and how to conduct peer review](#howpeerreview). The last chapter covers [problems with peer review](#problempeerreview) and this digs into some of the real biases that occur during peer review, and with the reviewers themselves. Peer review isn't a perfect system, but in order to get the most from it we need to understand the weaknesses, both as authors and as reviewers. Only through this understanding can we reinvent the publishing system. We should not expect to do away with peer review (although this has been suggested many times in the past, and no doubt will be suggested again in the future), but by understanding the the biases that exist we will be able to make sensible choices despite the limitation.

Changes in the peer review system can make the difference between a highly biased system where editors manipulate the content for their own purposes ([networks](#networks)), or a system that is fair and equitable to all.

## Impact factor
[Impact factor](#impactfactor) is a simple metric, and as such there is no need for it to be anything more. But the publishers have managed to use impact factor to their advantage such that it has become of overriding importance for many publishing in the biological sciences today. But it doesn't have to be this way. To make the most of the system, you will need a thorough understanding of the way in which impact factor can [control other aspects of publishing](#money), and how the behaviour of the [editor can have a profound effect on the impact factor](#increaseIF). The higher Impact Factor results in more submissions, and this in turn will mean that the editor will have more power over the content of their journal.

## Citations {#cite}
Sitting above peer review and Impact Factor in Figure \@ref(fig:publishing-triangle) are [citations](#citations). Compared to the other two parts of this [wicked problem](#wicked), citations seem to be blameless and without the potential biases of the others. However, citations are the units of control for Impact Factor and can be manipulated by both editors and peer reviewers. [Metrics](#citations) driven by citations are also at the heart of many of the problems in today's publishing world.

-----------------------------------------

The relationships between Citations, Peer review and Impact Factor are indicated with double headed arrows in Figure \@ref(fig:publishing-triangle). There follows an attempt to describe these subtle differences, and the way in which they impinge on the gatekeepers at the centre.

#### Citations and impact factor  
The more citations a journal gets (in the first 2 years following publication), the higher its Impact Factor. Higher Impact Factor journals will want any submissions that they choose to have a good chance of garnering as many or more citations than their current impact factor following publication. This can mean that if you have co-authors that are in active groups, publishing a lot, they are more likely to have their work accepted in higher IF journals (they are also likely to have more experience at science and publishing).

#### Impact Factor and Citations

The relationship flows in the other direction as the higher the impact factor of the journal that you publish in, the more likely your work will be seen, read and cited. Citations garner more citations, and your work bearing your name will become better known within your discipline impacting your chances of getting a job, being invited to give talks at conferences or at visits to departments.

#### Impact Factor and Peer Review

Your peer reviewers are likely to be directed differently depending on the Impact Factor of the journal. As the Impact Factor increases, so the Peer Reviewers will be asked about the novelty and impact (i.e. citations) of your work.

#### Peer Review and Impact Factors

The impact factor of the journal may well influence which peer reviewers are prepared to review your work. In theory, better read (usually more senior) reviewers will provide more insight and may well refuse to review for journals that they consider do not garner sufficiently interesting (novel and impactful) work.

#### Peer Review and Citations

Your reviewers are likely to look at the papers that you cite and this will inform them of the scholarly quality of your work. When you miss important citations, or omit contrasting viewpoints, you demonstrate a failure in your scholarly undertaking to read the literature.

#### Citations and Peer Review

Peer reviewers may well suggest citations for you to consider in your work. These suggestions can be legitimate attempts to increase the value of your work, and through these citations its visibility. There are also reviewers who will suggest that you cite their own publications, and even editors that suggest that you cite papers in their journal.

## Open Science - a vision of the future {#openscience}

The way I advocate in this book is toward a vision for a future of Open Science (Figure \@ref(fig:open-science)). This future is both open and transparent. The transparency means that there is no need for this book, as there will be no hidden agendas or assumed knowledge needed for Early Career Researchers. Instead, it will be a 'what you see is what you get' system.

(ref:open-science) **A simple schematic for an Open Science publishing model**. Open science relies on the open nature of publication, communication and data. The gatekeepers still lie at the heart of this model, but are principally involved in ensuring the open and free flow of scientific information. This ideal world is free of the metrics that have dogged research in the past leading it into a profit making game for private publishers, and promoting the careers of a few elite scientists.

```{r open-science, echo=FALSE, out.width = '85%', fig.cap="(ref:open-science)"}
 knitr::include_graphics("figures/open-science.png")
```

The simple schema for Open Science shown in Figure \@ref(fig:open-science) is modified from O'Carrroll et al. [-@ocarroll2017providing]. The three areas of Open Science start with Open Data, the need to share both data, the code to analyse data, and the details for open source software with which to do the analysis all within open data repositories. The sector on Open Communication replaces the current closed peer review system with an Open framework where all actors are named and any interests declared. These include any journal gatekeepers (if involved). Lastly the publishing of the work is Open Access for other scientists and the public. This includes both proposals, preprints and published articles.

At the heart of Open Science publishing is ownership of the content. Like many aspects of changing the model from closed to open, implementation of the Rights Retention Strategy [see @janickehinchliffe2021explaining], through mechanisms such as [CC-BY](https://creativecommons.org/licenses/by/4.0/), is available today. This is something that you, as an author, should insist on [anywhere you publish](#copyright). If the journal refuses, then **think about what are they saying to you.**

In my opinion, the [Open Science framework](https://osf.io/) is incompatible with the for-profit scientific publishing model that drives the current model (Figure \@ref(fig:publishing-triangle)). It can be hoped that the transition from Figure \@ref(fig:publishing-triangle) to Figure \@ref(fig:open-science) will be swift and happen within your career. However, the actors at play in this system are not neutral, and need both bottom up challenges (from yourselves as ECRs) as well as top down pressure (especially from large funding agencies). We should acknowledge that changing our research culture within science is not an immediate process, but requires a suite of cultural changes (Figure \@ref(fig:culture-change)), that will start with early adopters and ultimately end with regulated policy [@nosek2019strategy]. These changes will be as exciting as they are challenging, and I hope that the contents of this book will equip you to participate fully.

(ref:culture-change) **Changing the prevailing culture in science will take time and effort, but is currently possible**. This book attempts to provide information about how a change towards Open Science is currently both possible and easy (green), but requires widespread adoption among the biological sciences community in order to make it normative and rewarding (orange), until we reach ubiquity through policy (red) [Redrawn from @nosek2019strategy].

```{r culture-change, echo=FALSE, out.width = '85%', fig.cap="(ref:culture-change)"}
 knitr::include_graphics("figures/culture-change.png")
```