### 2021-09-05 2.05-openaccess.Rmd

# Open Access or a paywall for your manuscript? {#openaccess1}

There are many well established benefits to publishing an Open Access (OA) paper where there is no pay wall to any readers. These include, increased citations, increased exposure and coverage by the media, not to mention increased interactions with the public and the moral and ethical duty to be able to share research as widely as possible. Many studies are now suggesting that the advantages go even deeper [see a collection of studies by @tennant2017open]. However the current reality is that most publishers will then require you to pay in order to produce your publication open access. There are chapters later on in this book that discuss the importance of [open access](#openaccess). In this chapter, I review the different OA models available. At this period in time, the names are somewhat fluid, and you may not find the specific term mentioned here on the publishers’ website.

(ref:Paywall-graphic) **The world of Open Access (OA) is built on [shaky foundations](#openaccess).** Much more of journal content is still within the shadow of the paywall, even for journals that practice Hybrid and [Bronze](#bronze) OA. When you are inside your institution, you may not notice the paywall due to seamless integration with many literature databases. You will be far better acquainted with the paywall if you are from a smaller institution or trying to access literature from outside your institution. Once you are the wrong side of the paywall, your options for most content will be Grey and Black OA. There are regular assaults on these two initiatives by publishers.

```{r Paywall-graphic, echo=FALSE, out.width = '95%', fig.cap="(ref:Paywall-graphic)"}
 knitr::include_graphics("figures/Paywall-graphic.png")
```

## Closed Access: i.e. The Paywall {#closed}
This refers to the need for your institution or you personally to be subscribed to the journal in order to access the content. This is the traditional model in academic publishing. If you are a member of an academic society then you may get access to their society journals through your membership. This is still effectively a paywall that is maintained by the publisher and the society together (Figure \@ref(fig:Paywall-graphic)).

It is sometimes hard to know if there is a paywall if your university subscribes to the publisher or the journal that you are interested in. There has been a lot of headway made in having seamless integration and access to articles behind paywalls from within university IP addresses.

If you try working from home then you will quickly find out which journals exist behind paywalls. There is another chapter with ideas on how to [get around the paywall](#paywall) if you need to.

Although publishing an article behind a paywall is often frowned upon these days it usually means that there'll be no cost for you as the author, and so for many academics this is still the only real option in terms of publishing their scholarly work in an academic journal.

There follows a brief description of each of the OA models. See Piwowar et al. [-@piwowar2018state] for a historical review of these different models:

##Open Access 
There is a bewildering array of different Open Access types. Entries in the list below are not mutually exclusive. Moreover, you may not be aware how your manuscript may be treated, and it is not always possible to predict the level of OA your article will be published with in many journals. Each journal will likely have their own policy, tied to the publisher in many cases.

### Gratis OA {#gratis}
Gratis OA simply refers to the ability for anyone to access the text free, and without any [paywall](#paywall).

### Libre OA {#libre}
Some definitions of OA extend beyond simply being allowed to access the text for free ([Gratis OA](#gratis)) to it bearing a license so that you can share, redistribute and reuse (Libre OA). With Libre OA everyone is able to share, redistribute or copy the content. The Creative Commons (CC-BY 4.0) licence allows others to “remix, transform, and build upon the material for any purpose, even commercially”.

A good example of this is my ability to publish the following Figure \@ref(fig:Piwowar) from Piwowar et al. [-@piwowar2018state] as the content of _PeerJ_ is published under CC-BY licence.

(ref:Piwowar) **Reproduction of material is possible if it is published under CC-BY**. Figure 2 from Piwowar et al. [-@piwowar2018state] shows the increasing number (A) and proportion (B) of a random sample of 100 000 papers published with a CrossRef DOI with an increasing trend in Gold and Hybrid OA publications since 2005.

```{r Piwowar, echo=FALSE,  out.width = '95%', fig.cap="(ref:Piwowar)"}
 knitr::include_graphics("figures/Piwowar.png")
```

### Green OA {#green}
This is an option no matter what journal you publish in, including [closed journals](#closed) behind the paywall. The accepted manuscript, before it has been typeset, is deposited in your institutional repository; usually your library hosts this. This is sometimes referred to as a **postprint** (manuscript with changes made following peer review) - the same version that has been accepted by the journal. Once your paper is published by the journal anyone can access the manuscript that was accepted because it is available free from your institutional repository. However the typeset paper will remain behind the publishers’ paywall (i.e. [closed](#closed)). Note that this version will not be the [Version of Record](#VoR), it may be that some of the text will change with [copy editing](#proofs). However, for people who want Open Access your work, this Green OA version should suffice.

When you deposit your author version to your institutional repository, make sure that you add a [Creative Commons Attribution](https://creativecommons.org/) license (CC BY). In theory, this turns your work into something that is [Libre](#libre) Open Access. Although the wording will be the same in the [Closed OA](#closed) version eventually available from the publisher, the prior attribution of CC BY to your Green OA work should mean that all subsequent versions are [Libre](#libre) OA. There are a number of reasons why this may not be the case, and you should be aware of two of these [@khoo2021why]. First are your obligations to your funders. For example, your funders may have specific requirements about how you publish, and prior rights attribution may conflict. Funders may have other restrictions on your publishing including embargo periods. With many publishers you may be required to sign an acceptance letter that specifically states that it's supersedes any previous licensing commitments including self archiving copies that you make CC BY. Your previous CC BY commitments to your pre-print may then place you in a legal conflict. Publishers are unlikely to assert their license unless you try to assert yours.

If you are unsure about whether or not to deposit a copy in your institutional repository and what license should be attributed to this, then ask your librarian. You may need to furnish them with details of your funding and potential target journals where you'd like to publish.

### Bronze OA {#bronze}
Bronze Open Access is an option at the discretion of the publishers. Only some publishers do this for some of their content. For example, they may decide that a certain thematic issue should be open access, an editorial, or review article. There is some debate about whether this bronze model is truly OA as it often still carries copyright restrictions.

### Delayed OA {#delayed}
Some journals choose to have their archived content available for all readers without a paywall. This could be after two or five years. Like [Bronze OA](#bronze), there is some debate about whether delayed OA is truly open as it often still carries the publishers’ copyright restrictions.

### Gold OA {#gold}
In a journal with gold open access it is compulsory for you to pay the open access fee (or Article Processing Charge, often referred to as APC) in order to publish your paper. These APCs can be very large, often more than \$1000 (see [Part IV](#money)). If your choice of journal is gold open access then make sure you know where the fees are coming from. The history of Gold OA is an interesting one that you can read in [Part IV](#PLOS).

Examples of gold open access journals are _PLOS ONE_, _F1000Research_ and _PeerJ_, and an increasing hoard of [no impact factor journals](#noIF). The advantage to these journals is that as soon as you publish your work in them, everyone will have access to it without any paywall. The disadvantage is that you may not be able to afford to publish there. The disadvantages of these journals are now becoming very prominent in the Biological Sciences, such that you may be excluded from increasing numbers of journals in your field. Alarmingly, journals with higher impact factors are also charging ever increasing sums to publish Gold OA [@gray2020sorry]. Gray [-@gray2020sorry] makes the important point that prestige (often confused with [Impact Factor](#impactfactor)) is being allocated a higher price in Gold OA, that is likely to disadvantage and disenfranchise scientists from less wealthy institutions and countries. This is likely to reinforce an increasing dichotomy between rich and poor researchers. We will take another look at who pays for OA [later in the book](#openaccess).

Watch out for predatory journals among those journals with Gold OA. Predatory publication will be [discussed in full later](#predatory)).

### Hybrid OA
Hybrid OA journals are increasingly the norm. You can decide upon acceptance of your manuscript whether or not you want to pay the fee to make your article open access. Again, note that your institution may have a deal with the publisher that means that anything you publish there is open access. It is well worth knowing these things in advance before you submit. If you can't afford to pay the open access fee then your manuscript will remain behind a paywall and be only available to subscribers.

A slight variation on hybrid OA is when you are a member of an academic society that allows its members to publish open access without extra payment. As a student you are likely to get very discounted membership to an academic society Which might make it very cheap for you to publish open access with them.

### Platinum or Diamond OA {#diamond}
This is without doubt the best OA model, and the one that we should all strive for. In a platinum or diamond open access journal you do not have to pay any money but everything that is published is open access. In order to do this these journals are often subsidised by governmental or philanthropic agencies. Some university presses are also in the habit of publishing platinum or diamond OA when it meets with their stated mission. These journals are very rare but they do exist, and it is well worth looking for them.

Another great new model that is diamond OA is the concept of [Overlay Journals](#overlay). Although there aren't any overlay journals in the biological sciences at the time of writing, I think that it's only a matter of time.

### Black OA
This refers to the placement of published material onto a pirate website such as [Sci-Hub](https://sci-hub.se/). Sci-Hub is considered by most governments to be illegal and may be blocked by your institution or country. However they do a great job in making scientific publications open access and need to be supported for that. Many scientists all over the world depend on Sci-Hub in order to access literature and therefore conduct research. In addition, there are a number of other Black OA sites: [libgen](http://libgen.li/),  [Unpaywall.org](https://unpaywall.org/), [Open Access Button](https://openaccessbutton.org/), [freefullpdf](http://www.freefullpdf.com/).

### Grey OA
You can find grey OA repositories of published material on Academic Social Network sites in which you need membership to access such as [ResearchGate](https://www.researchgate.net/) and [Academia.edu](https://www.academia.edu/). The legality of such sites is regularly questioned [see @piwowar2018state for more details]. There has been legal action with thousands of members being issues with take-down notices.

## Supplementary Information and Data
Whichever route you decide to go for your manuscript, please do not place your data with the publisher. There are some examples where publishers choose to place both data and supplementary information deposited with them behind a paywall, even if the article is available Open Access. We also need to ask whether the publisher has the long term vision to curate data, especially when the expense associated with this will rise over time as datasets accumulate behind their paywall. Elsewhere in this book you will find some suggestions about [what to do with your data](#transparency) to make it available for all.

## Unsure what you can legally do with a published paper?
If you don’t know what level of copyright exists on something that you have published, then you can find an aggregated set of publisher policies at [Sherpa Romeo](https://v2.sherpa.ac.uk/romeo/). This is a really nice database which provides a very simple summary by journal. You can also use this to check out a journal that you are thinking of publishing with. If you are still in doubt, then consult your librarian.