### 2021-09-05 _404.Rmd

# Page not found

**How to Publish in Biological Sciences**

The page you requested cannot be found (perhaps it was moved or renamed).

You may want to try searching to find the page's new location, or use the table of contents to find the page you are looking for.

**To continue, you can simply click on the contents (left),  or click on the image (below) to take you back to the beginning of the book.**

```{r fig.align='centre', echo=FALSE, include=identical(knitr:::pandoc_to(), 'html'), fig.link='https://johnmeasey.github.io/How-to-publish-in-Biological-Sciences/'}
knitr::include_graphics('figures/cover.jpg', dpi = NA)
```