### 2021-09-05 3.01-Title.Rmd

# Title page {#title}

Most of the following information in this section on the title pertain to getting a manuscript ready for submission to a journal. For your thesis, I would recommend that you have a title page for each chapter that lists any collaborators and their contributions so that this is transparent for the examiners. It also helps break up the document and can be a good spot for a non-academic illustration (if your institution permits).

Nearly all journals will require you to have a title page for your manuscript. This may or may not include authors. Check to see whether
the journal that you are submitting to conducts double-blind review. If they do, they will require a title page without any indication of authors or institutions (they will also likely ask you to remove the acknowledgements section). This should be clearly specified in the instructions to authors of your chosen journal.

### Names and addresses are important {#ORCID}

Getting authors' names correct, their correct addresses (many people have more than one affiliation), can be tricky. Check that you have all of the information required for a title page before you submit. While you need the names of your co-authors to be displayed correctly, you do not need their titles (no Dr. or Prof., etc.).

Indicate clearly who is the corresponding author. Normally, if you have done all of the rest of the work, the corresponding author should be you. You need to learn how to start this role at some point, so it might as well be now. Having the correct name and address will be important to each author. Along with the author names and addresses you should record the ORCID of each author. The ORCID is simply a unique identification code for individual researchers. ORCID is a non-profit organisation and there's nothing sinister in signing up. If they (or you) don't have one, then you should ask them to create one before the submission. It takes less than 5 minutes. Go to www.orcid.org

## The title

The title of your paper, chapter or book is the first thing that any reader will read, and so should be well considered. If your reader
cannot quickly understand your title, the chances are that they won't bother reading any further. Your title will be your selling point, and your aim is to use it to draw your readership in. Once you've managed to inveigle your readers to download your paper, your title is also their hook for remembering your paper in their database of thousands of others. Having their key-word in your title will help here, and as ever with writing your challenge is to think like your reader. The best titles are those that sum up the entire study in five to seven words. This is best done in a narrative that tells the story ([see part 2](#storytelling)) of your manuscript in its entirety. This may sound daunting, but you should get into the habit of summing up your story quickly (for friends, relatives and work colleagues). Then it's a question of refining this story into the short single sentence that makes up the title. While the narrative approach may not work for you, you do want the title to provide enough information so that the potential reader knows what they will find before they open it. Your title doesn't just have to work for you, it needs to work for a wide audience.

Some people are excellent at writing titles that contain puns of well-known phrases or sayings. These can be brilliant, working both to
inform what's in the paper as well as providing some familiar input that helps retain them in memory. However, many fail to do either and are simply a waste of space. If you are tempted to use a pun as your title, make sure that it is widely appreciated, and not just among your co-authors and lab.

You don't have to come up with a killer title from day one. Most of my manuscripts have a working title that gets revised as I write, and is always open for change before submission. If you have great ideas for a title, do note them down. I find that the more options I have, the more likely I am to come up with something that works for everyone. It also helps to mix and match from a set of candidate titles. Once you have come up with something that looks good to you and your colleagues, test it by entering it into your database of choice (with the default being Google Scholar). Your first 10 results should include a set of papers that you have likely cited in the upper area of your introduction. If you don't recognise any out of the top 10, it's time to look at another of your candidate titles.

### Some title ideas to start you off

-   Don't start by looking for the best title

    -   Write a number of candidate titles and ask your co-authors to vote for their favourites.

-   The shorter and catchier your title can be the better: 6 words (±1) is an ideal.

    -   Allow yourself a longer subtitle if needed, but don't go over 20 words total (some journals may limit your total to less).
    -   Consider the (former) 120 character limit of a Tweet as an upper limit

-   Do include your principal finding if possible

-   Include as many key-words as you can

### Things to avoid in your title

-   Don't feel obliged to include [taxonomic terms](#taxonomy) unless it
    is relevant or compulsory

    -   Some (taxonomically minded) journals will insist on the species
        name followed by the taxonomic authority, and/or the family and
        order, in your title

-   Avoid obscure or specialist words that won't be understood by your
    readership

    -   There are times when key words are necessarily specialised and
        your readership will expect this, but simpler words in your
        title will open up your readership which will otherwise remain
        narrow

-   Don't simply define the scope of your work without including your
    content.

-   Don't have your title as a question. Rather provide the answer!

## Key-words

The key-words are a way for readers to find your content with searches. Typically, the advice is to use words that are not in the title or abstract. This is because many databases have combined searching facilities for title, abstract and key-words. I struggle to think of
appropriate key-words, and so make a list of some of the [big idea](#bigidea) words and short phrases from the introduction. Then I
tend to look at articles in the same genre, and see what key-words they have used. As with the title, I suggest that you enter your chosen key-words into your literature database of choice [see part 2](#database) and see what comes back. You should see a group of papers that look wholly familiar and preferably those that are already cited in your manuscript. If not, it's time to look again.

### Selecting appropriate key-words

Key-words are very useful in your studies, because if you have a good
selection, they can help nail down a good proportion of the literature
that you will need to read during your studies. Moreover, if you have
the best selection of key-words, then you can set up some automated
notifications for when new items are published. The only problem then is
pulling together the correct key-words.

When you submit a conference abstract, or a paper for publication, you
will also be faced by a demand for key-words. This is when I often draw
a blank. What will be the appropriate key-words for my study? I can
often think of one or two, but regularly journals want at least five.

A method for finding key-words Here is my method for finding appropriate
key-words. It's quick, but it does require a piece of free-ware:
[VOSviewer](https://www.vosviewer.com/), which for me has become an invaluable research tool.

-   Go to your literature database of choice (i.e. Web of Science,
    Scopus, PubMed, Dimensions.)

-   Search for documents using the key-words that you are sure are
    appropriate

-   Constrain the results to \~500 documents

    -   You can either do this with the search date (e.g. the most
        recent 5 years)

    -   Or you can simply take the first 500 documents that are found

-   Download these \~500 documents in a tab delimited text file. In
    VOSviewer Press the 'create' button

-   Create a map based on bibliographic data

-   Read data from bibliographic database files

-   Select the file that you downloaded in the appropriate tab

-   Type of analysis

    -   Co-occurrence

    -   All keywords (you can choose here)

    -   Full counting

-   You should see the total number of key-words in your downloaded file
    now

    -   Change the minimum number of occurrences of a key-word so that
        you have \~100 results (again you can choose what suits you)

-   Press finish

You should end up with a network like the one below. Here I have used the key-words "invasive" and "fish" (Figure \@ref(fig:key-word-selection)).

(ref:key-word-selection) **A network of key-words for "invasive" and "fish".** This network represents key-words that are repeated 10 times or more in the Web of Science together with "invasive" and "fish".

```{r key-word-selection, echo=FALSE, fig.align='center', out.width='90%', fig.cap="(ref:key-word-selection)"}
knitr::include_graphics('figures/key-word-selection.png')
```

In VOSviewer, you can highlight any one of these key-words and see exactly what combination they have been used in. I have selected
examples that occur 10 or more times in my downloaded file. This means that I can be fairly sure that these are relatively common key-words to use in combination with the ones I know are good.

The larger the panel in this network, the more frequently the word is used (see "invasive species" in Figure \@ref(fig:key-word-selection)). This should help you when you select your own key-words. For example, even though I had used the keywords "invasive" and "fish" to generate this network, one of the first things I noticed is that the term "invasive species" is far more common than "invasive". Hence, the first thing I should do is to change the first of my key-words.

You should note in Figure \@ref(fig:key-word-selection) that some of the key-words relate to specific taxonomic groups. Others include the habitat in which the fish were sampled. Now I have a shortlist from which to pick the remaining 5 key-words that I need in order to submit my abstract or manuscript. Once you've made your selection, you can go back to your literature database, add this combination of key-words into a search and see what comes out. If you've done it right, you should see some familiar papers on similar topics to your own.