### 2024-03-15 2.17a-courses.Rmd

# Choosing your courses

As a postgraduate student, your university should offer you a number of courses designed for postgraduate students. In many countries, these postgraduate courses will be compulsory and admission to the graduate school (Admission to Candidacy) will only be made after you have completed sufficient courses with grades to qualify you. For example, in the USA there are normally two years of these postgraduate courses followed by three years of working on your thesis project.

Given that this is actually a really long time, it's worth thinking about how you can get the most out of the courses that you choose, and how best to use this time in general. In this chapter, I will not tell you what courses to take, but try to give you some insights into how to pick the courses that you are going to get most use from, both during your PhD studies and in your life thereafter.

## How to choose your courses

1. There will be some courses that align directly to the contents of your thesis and you should do them
1. There should be a [writing course](https://johnmeasey.github.io/Thesis-Writing-Guide/) (and if not you can follow a link to my own) that you should attend as everyone needs help with writing
1. A time management course would also be very helpful, if your university has one
1. Look back at your short and medium term goals (see [here](#Plan)) and consider which courses could be relevant in your future
1. Do some courses just because you are interested. Given that you have already done a lot of courses at university, you are going to need some that hold genuine interest for you. Remember that diversity is an important aspect of science, so think as widely as possible.

### Attending a postgraduate course at another institution

Some universities allow their students to attend postgraduate courses at other (usually nearby and affilitated) universities. If you know of a course that it being conducted at another university, but that you would like to attend, it might be worth approaching your postgraduate teaching co-ordinator to ask them if you can go.

There is also the opportunity to conduct a postgraduate course online. Again if you know of one, or put in the time to find an appropriate one, it may be worth approaching your postgraduate teaching co-ordinator to ask if you can do it.

The most important consideration is whether or not your institution is able to use marks generated at another institution in their own system to assess your postgraduate courses. Even if they don't allow you, it's worth talking to them about it as it may help those that come after you to find the course that they want. Or it might help stimulate to offer something similar at your institution in future.

## Due diligence during courses

You need to make sure that you get the most out of these courses.

Looking back at being an undergraduate, you may have been focussed on grades or shortcutting work on courses, but for postgraduates this must have changed.

## Getting the most our of your courses

Remember that it is not direct applicability that is of the only importance.

## Building your network

Courses are designed for postgraduates and usually taught by the most experienced professors. Use your time with them to [increase your network](https://howtopublishscience.org/advisorless.html#networks).

In addition to the professors that are giving your courses, you will also learn about other academics at different institutions. While these people may be far from you, their work can be very relevant to your own. Using your time wisely on your course also involves reading work by these people. If your advisor agrees, then it may be possible for you to approach some of these top researchers to participate on your committee.

## Time off from courses

As you are in the biological sciences, there may be good reason why you need to go on a trip and not attend part of a course. For example, your advisor may ask you to participate in lab fieldwork that is relevant to your project. You may want to attend a conference to present the results of your MSc work. You may have any number of private or professional reasons why you need to miss lectures or parts of your course.

 - Inform the course co-ordinator. Your institution may have a system where you require to be signed out of the class. Make sure that the course co-ordinator knows, and the lecturer of the missed classes (if this is a different person).
 - Catch up. Be diligent and catch up with all of the missed course work. 
 - Try to miss as little as possible. Taking your courses seriously means missing as little as possible.

## Make the most out of learning

I you haven't already worked this out for yourself, learning is a real gift and the postgraduate courses you are offered at universities are probably the ultimate in their level of sophistication and knowledge transfer. Take this opportunity seriously and you should find it extremely rewarding.