### 2024-09-11 1.04-email.Rmd

# Communicating by email and messaging apps {#email}

Throughout your PhD, you will be expected to communicate with your advisor and other professionals inside and outside of academia using email. In some contexts, you may also be expected to use social media apps like WeChat and WhatsApp to communicate either directly or in groups. The chances are that you have already written a great deal of emails and instant messages in the past. However, this does not mean that the style and content of those emails and app posts will be appropriate when communicating during your postgraduate studies.

In this chapter I will start by considering some issues around correct use of email, and then move onto direct messaging apps [below](#wechat).

## Email

In addition to meeting in person, email has become a standard way of communicating between students and their advisors. If your project has any collaborators, then you will likely also interact with them over email, perhaps without ever meeting them in person. Although other platforms for communication are available (e.g. Slack), email has the advantage that it provides a paper trail (time stamped, and institutionally traceable), is rapid, can include attachments and can include additional advisors or collaborators.

Email is a very low-cost method of communication, but it isn’t free. Every email sent takes energy to send it, and the bigger the size (especially sizes of attachments), the more electricity is used to send them through your institutional server, or beyond. This is actually a reason why you should not be sending superfluous emails as the bigger they are, the more energy they use.

An important point about emails is that they can be kept and/or forwarded to third parties. This means that you are best advised not to include any content that might be considered (by anyone) as risky or contentious. In particular, avoid anything written about any third party, unless you are happy for them to see what you have written. There is also a legal implication in that they can be used in court, for example, if you have written something libellous about a third party. This is true even if you didn't write the email but simply forwarded the content that someone else wrote. Remember that what is funny to you might be offensive to someone else.

### The subject
Do include a succinct subject for your email. Do not leave this blank. Remember that your advisor, and other professionals that you communicate with, likely receives many emails in a day. They may not be able to respond to you immediately, but may want to know what the email is about before they open it.

The subject is especially important if you need something quickly (e.g. something signed, or for permission). If this is urgent, you might consider saying when the deadline is in the subject line. Obviously, you should avoid making urgent demands from anyone in your professional world, but sometimes this is necessary and everyone will understand. However, if you fail to communicate this in the subject of your email, don’t be surprised if your request is sidelined until after it’s too late.

### Writing style
In general, it’s best to keep your style brief. If you feel that you need to write a long email, consider ways in which you could better communicate the same information in a meeting, or in responses to comments in a chapter draft. I would discourage you from writing long emails as reading them will tend to take up more of your advisor’s time than other forms of communication.

I used to get annoyed with people who insist on starting every email with “I hope this email finds you well” or some such banality. I now understand that this is culturally important for some people who are used to using longer greeting formats and who are genuinely sending a message of well wishing. Some time back, there was an administrator in our small unit who used this in every email that they sent, irrespective of whether we'd just been talking over a coffee or had spent the entire morning in a meeting together. Hence, this overuse of something that could otherwise have been nice to read became irritating. My advice would be that if you do have a personal connection with someone and want to make this clear, then by all means do so. Try to make it personal and not generic. Certainly, refrain from excessive use where any phrase could become meaningless or irritating

There are whole books on how to write emails [e.g. @taylor2009email], so I'm not going to go into too much detail here. Suffice it to say that just like any other form of communication:

 - try to make the reason why you are communicating clear
 - read through what you've written before you press send
 - try not to be overly familiar or informal [@lewin-jones2014understanding]
 - be polite, as this may make a difference to you getting a positive response [@bolkan2012you]
 - let brevity be your watchword

#### Don't use slang or SMS language
I remember an email that a colleague received (many years ago) asking for a bursary. I don't remember the exact wording, but it went something like this:

> "Hey Nanni [her family name] - U gonna give me 1 bursary? I is the best student."

Needless to say, this student didn't receive a bursary, although they could well have been the best student in their class and fully deserving. Their skills at communication were not sufficient to warrant another look. And while this example may seem ludicrous, it appears that many undergraduate students do correspond with their lecturers in this way [see @stephens2009r].

As with all communication, you need to be focussed more on the recipient and their ability to comprehend the message that you intend. Although I will reiterate that you must keep your emails short, this should not be at the expense of comprehension.

#### Formalities {#formal}
The level of formality that you use will largely depend on your existing relationship with your correspondent. The formal tone is set immediately with the greeting that you use.

I am happy for people to start any email with “Hi”, but other advisors might prefer something else. If you are unsure about your advisor, then ask them. As a default, you should use the same style that they are using to address you.

When emailing someone that you don't know, use the most formal style: "Dear Dr. Blogs" or "Dear Prof. Bliggs". If you don't know what their rank is, use an appropriate level that you consider likely. Remember that there are conventions in different places (e.g. every academic in the US is a Prof.; in Europe Prof. is reserved for the highest rank only, and everyone else is Dr.). Similarly, when you [sign off](#signoff), use a formal [sign off](#signoff). As a general rule of thumb, you can use how they [sign off](#signoff) their email to address them. For example, if I [sign off](#signoff) an email "With best wishes, John", I expect that you will reply with Dear John or Hi John. However, if I [sign off](#signoff) "With best wishes, John Measey", then I am looking for a more formal response starting with "Dear Dr. Measey".

Remember that if you cc in others, you should move your formality accordingly. Bear in mind that emails can be kept or forwarded without your knowledge. You do not want to regret a particularly lax use of formality at a later date. I would suggest that the more people that you address your email to, the more formal you'll need to be.

In general, emails that you write as a postgraduate student are professional communications, so keep them as such. This does not mean that you need to start using citations in your emails, or that you can't use contractions (like can't, won't or didn't). You emails don't need to be formal as in the style of a published paper. As with the initial greeting, you should look at emails you receive to see how to set the formality of the tone that you will write.

### Why exactly are you communicating?
The first two or three sentences of your email should state why you are communicating and whether or not you are expecting a reply. Emails can be just for information, keeping someone up to date on your progress or letting them know that something they asked for has been done. If this is the case, then start your email with a statement letting the recipient know the reason why you are keeping them informed, and that you don’t expect a reply.

If you require a response from your recipient concerning a particular question, then you should flag this in the first line of your email. If their response is time sensitive (i.e. you need a reply before a certain date) then state exactly when this is. In the subsequent information that you write in your email, you need to make the question or questions clear.

If you need to write a long explanation in your email, then you might want to draw attention to the question that you want answering by making it __bold__. Another useful means of concentrating your recipients attention is to use structure.

#### Using Structure
I like it when people use some structure in their email. If it’s going to be a long email (which I’d advise against) then do include a short summary explanation at the start so that the reader knows what the email is about (especially if the subject line doesn’t fully explain). Using bullet points or numbered points in the email is also a great way of using structure, especially when you expect a reply on each point.

##### Replying inline
This simply means that the previous email is reproduced when you hit reply, often with some marker for what was written originally. I like inline replies as it’s simple to see that each point has been addressed. It’s also a useful way to be reminded of what was originally written.

Some email chains start to become a bit tedious when many inline responses make them so cluttered that it’s hard to see what’s going on any more. Use your common sense to decide when it’s enough.

##### HTML emails or plain text
Most email software now defaults to emails written in HTML, which means that you can use **bold**, *italics*, bullet points and most other features that you would in a word processing package. Some people prefer plain text only, but HTML is becoming more standard.

##### Email chain
Most email software has an option to include the original message in the reply. This is quite useful in that it can remind everyone about what was asked. As these kinds of conversations get longer, they can become very large and unwieldy.

##### Bullet pointed or numbered lists

I think that both of these are appropriate when you have a number of points that you’d like addressed.

#### Use of all Capital Letters - DON’T DO THIS!!!!

When reading any email when parts are written in All Caps, this comes across as someone who is shouting (see use in subheading). Not something that you should be doing in any professional communication. If you need to, then use bold or italics to emphasise points.

Similarly, I’d appeal for you not to use excessive exclamation points, slang (especially swearing) or other such frivolities in your professional email. It’s simply not a way of communication that you should be doing at work. Write emails to your friends like this if you want to, but I’d suggest that you do this from a private email account (and not your institutional address).

### Signing off {#signoff}

How is it best to sign off? Most professional emails sign off with one of the following:

 - With best wishes
 - Kind regards
 - All the best

Any of these is fine to use, as are the more formal classic letter writing “Yours faithfully” or “Yours sincerely”.

I quite like it when people sign off using their own language, as it (somehow) makes it more personal and sincere. I have worked in a number of countries around the world, and when there tried to use local language words to start and end emails.

### Who else should your email be sent to?

You may need to send emails to more than one recipient. If you are planning to send an email to a lot of recipients, then it might be worth checking this with your advisor first. You will also need to decide whether the recipients should be listed in the **To** box, CC or BCC.

If the email is sent to someone, then their name should be listed in the initial greeting: "Dear Drs. Bliggs and Blogs"

### When to CC someone

Only Carbon Copy (CC) someone if there is a real need for them to be included on the email. In reality, multiple addresses can be added in the To field, and anyone that the email is written to should be listed there.

In the business world, a lot of email correspondence includes the entire team in CC. This is usually not needed for academia. If you are in any doubt about whether or not to include someone in CC, then ask your advisor.

If you do put someone additional in CC, then it is polite to note this in the email. For example, in the initial salutation you could write “Dear Sue and Bob (and Warren in CC)...” I like to indicate the reason for including someone in CC at the appropriate point in the body of the email. This lets everyone know exactly why they have been CC’d.

> “Dear Sue and Bob, I am writing regarding the delivery of the 33 g steel ball that Warren (in CC) and I ordered on 12th November this year.”

Or

> "Dear Drs. Bliggs and Blogs (Prof. Blab in cc)"

Remember that someone in CC will be in CC again if the person you are writing to hits [Reply All](#replyall).

#### Use of BCC

Blind Carbon Copy (BCC) is a very useful way of copying in another person without the original person or anyone in CC knowing. But you should use this feature of email with extreme care. Many people who receive an email in BCC are not immediately aware, and it has happened to me that people I have BCC’d have then hit “[Reply All](#replyall)” and immediately disclosed their presence on the email. This is very embarrassing. Thus I would urge you to only use this when the person you intend to BCC is aware of the fact.

A common use of BCC is when emails are sent to a great number of people and to avoid having anyone use [Reply All](#replyall). If you have a great number of collaborators (for example on a literature review or a large author list), then you can address the email to yourself (or your advisor) and BCC everyone else.

Note that if someone is on BCC, then they won’t get included when any of the rest of the group (addressee and CC) hit [Reply All](#replyall).

### Reply or Reply All? {#replyall}

Hitting Reply All instead of Reply is perhaps one of the most common mistakes that I see with email. Some email software has Reply All as the default option whenever you hit reply. If yours does this, I’d urge you to change it immediately. The use of emails to large numbers of recipients is common (especially in companies), but it gets extremely annoying when half of this list hits Reply All just to confirm receipt or even worse give some inane chatty response.

If you are going to use Reply All, before you send the email check:

 - that it is appropriate everyone on the list receives your reply
 - that there is not inappropriate information in the email chain
 - that you have added the correct attachment

If you have an email inbox that serves multiple email addresses, then it is worth noting that when you hit Reply or Reply All, the default is for it to send from the address that the email was sent to. You can change this default, or you may want to change from (for example) gmail to your institutional address.

### Institutional or personal email addresses

As a student at your institution, you will likely be allocated an institutional email address. You probably also have your own private email addresses (e.g. Gmail, Hotmail, Yahoo, etc.). Some institutions have great problems with their mail servers, and many academics will abandon their use of their institutional email address because it causes too much trouble for them.

It is worth asking your advisor which address they would like you to use (if you know that they have several). However, I would suggest that you make use of the institutional address for all formal communication so that you have a record of [due diligence](#deliverables) should you need it.

#### High Importance
Some email software has the ability to flag an email as important. Avoid using this unless you have something with real importance, like some decision or response needed that same day. Otherwise, try not to use this feature.

#### Urgent emails {#urgent}
Sometimes it is necessary to communicate urgently. Deadlines are unforgiving and your advisor and other people are not always aware of when they are or what they mean to you. Deadlines are also very stressful, and you might be more prone to miscommunicating or making mistakes.

If you need to communicate urgently by email, then here are some tips about how to get a swift response:

1. **Urgent for whom?** I think it's worth starting by asking yourself why exactly this message is urgent, and for whom. It could be that this is only urgent for you and your deadline, or it could be that this is a shared urgency among you and the people you are corresponding with. Alternatively, the urgency might not be for you and only for the other person/people. Starting out knowing this information is important as it will help you set the tone for the message. 
2. **State that it is urgent in the subject**: Start the subject with "Urgent" and give a timeline by when you need an answer in the subject line. Most people will receive their emails on their phones and at some point they will see the subject line without reading the contents. If the subject line can inform them of what the topic is and when the urgency is for (e.g., Urgent: signature needed in the next 2 hours, the next 2 days, the next 2 weeks, etc.), this will really help them knowing exactly what is needed.
3. **Within the message**, provide the explanation again together with a clear and concise account of exactly what needs to be done. If you need a signature, then remember the attachment that needs to be signed. The more concise you can be about what is needed, the faster it is likely to be done.
4. **Follow up with an instant message or text**. If this is urgent and especially if it is outside office hours, then follow up with an instant message to check that they have received your message. See [section on texting below](#WeChat).

Before you send the email, take a few moments to read it through and check that everything is correct. If it really is urgent, then the last thing you need is to have some to and fro over the actual meaning of what is written.

Some deadlines are really important, and some correspondents are famously late with responses. In this case, consider providing a deadline one or two days earlier than the actual deadline. This only works when you have a couple of weeks to go, but it might help you build in more time and reduce the urgency. For example, you can use the [Send Later](#later) feature to send a couple of reminders building up the the date. But remember to cancel them if you get what you need before!

#### Sending emails outside office hours
Personally, I have no problem with people sending emails whenever they want, as long as they don't expect replies to come out of office hours. I have often noted that chapter drafts appear in my inbox late on a Friday night. My suggestion to students is that this is not a good use of their time. Rather keep the draft and read it later on Sunday evening themselves after having left it alone for the weekend. You are more likely to spot errors and improve your own writing this way. It will certainly be better than it sitting untouched in my inbox all weekend.

If you don't know whether or not you will get back to your draft on Sunday evening, this would be a great time to use the [send later](#later) feature.

#### Send Later {#later}

This is a new feature on many pieces of email software that I really like. Essentially, it allows you to compose an email and stipulate the time and date that you want it to be sent. I find this very handy if I know that someone is on leave. I'd prefer that my email not sit in their backlog, but arrive once they are back from being away and have already cleared their inbox.

Another scenario that I use for send later is [when I have a deadline](#urgent), but am awaiting feedback. I can make sure that I meet the deadline by sending the email with report or manuscript attachment or whatever it is that I have to send before the deadline. However, if at the 11th hour I get the feedback from the third party that I'm hoping to include, I can still find the email with report (it's usually in the Draft folder), make any amendments and then send it off.

#### Forwarding emails {#fwd}

One of the most common ways of emails being inappropriately shared is when they are forwarded, or additional people are added to a reply much later in an email chain. I have received many emails where the sender was clearly unaware of what the chain contained.

In general, it’s not often appropriate to forward emails to others.

### Out of Office autoreply
I find the autoreply saying that someone is away very useful. It lets you know whether or not to expect a reply, and might mean that you then write to someone else instead. If setting your own Out of Office autoreply:
 - Do include the date when you will be back (and check that it’s correct when you send it)
 - Do state whether you are on holiday or in the field 
 - If there is another means of reaching you (when you aren’t on holiday) then give this

 Becase you don't know who will receive your autoreply, make it [formal](#formal) and polite.

### Using email signatures and banners

An email signature is a useful way of explaining who you are at the end of your email. I find them very useful sources of information like websites, telephone numbers, physical locations (i.e. address). They can also include more information. For example, my signature currently contains a list of journals that I edit for and links to this (and other) books. For this reason, signatures are useful, especially on first contact. Thereafter, signatures don’t really need to be sent with replies and in email chains. Most email software contains options for when signatures are sent.

Some institutions will automatically add a pictorial banner to the bottom of every email sent, advertising how great they are. Others will add long legal statements with disclaimers. Obviously, you will have little power over whether or not these are included by your institution. However, some students create and add their own banners to their emails. I would discourage this simply because it takes more energy to send, and uses up considerably more space than text only signatures.

### Using emails as a record of due diligence {#deliverables}

For students that are having problems with their advisor, especially with lack of communication, or non-communication, email provides a very important record to document your due diligence, as well as being able to document your advisor’s lack of timely response (hopefully you already have a student-advisor agreement to refer to). Make sure that you use your institutional email address and the institutional address of your advisor. If they prefer another address, use this in cc in addition to their institutional address. The reason for this is that if you are going to use this paper trail to document your communication, you will need to produce these emails for your institution and they may only consider emails that have been sent using their services.

If you suspect that you may need to use your emails in this way in the future, it is worth re-reading them, prior to sending them, and think about how they could be interpreted by a third party.

**If you are communicating with your advisor on a web based platform (like Slack, WhatsAPp or WeChat), then do you need to send them emails as well?**

In most institutions, postgraduate students and advisors sign agreements before they start working together. These agreements are likely to come along with deliverables that may include progress reports and chapter drafts. It is well worth looking at this agreement and see what you've signed up to. I would maintain that at any point that your agreement states that you need to 'communicate' with your advisor, this should be done via email. Similarly, if the agreement talks about you needing to send thesis drafts, do these via email. As stated above, emails are useful because they can be traced, and institutional emails especially so. If either party claims that they 'did no receive' an email, this can be independently verified. If you sent something via Slack and at some point your advisor claimed that you didn't, how do you verify this later?

Of course, everyone hopes that there won't be a breakdown in communication during your studies. However, it's worth starting your professional career as you mean to go on.

### Keeping emails {#keepemails}

There is a vast array of options when it comes to keeping emails. I know people who consistently keep an empty inbox and do not keep anything. I know other people who keep everything. Some people put every email they received (or send) into a folder. Others keep them all in the inbox (ok - I'm the only person I know who does this).

Many institutional email servers will have a limit of what you can keep. These days it can be quite large (e.g. 1 TB). So should you keep all correspondence? I would say that it is worth keeping anything that relates to [deliverables](#deliverables) that you have.

### Institutional Policies

Your institution is very likely to have a policy on email use. If you are unsure and there is no summary available to you from your department, you can ask for a copy of the policy to see what is officially allowed and what is prohibited.

#### Folders

I like to use tags to group particular subject areas together. I see no point in using folders for people as any search for a person can quickly recover all emails from someone.

I find that after 2 years, most correspondence is no longer useful. There is the odd occasion when I delve further back in time, but older institutional emails get archived after 2 years.

## Use of WeChat or WhatsApp during your PhD studies {#wechat}

[WeChat](https://wechat.com/) and [WhatsApp](https://whatsapp.com/) are two social media direct messaging apps that currently dominate the markets in Southeast Asia and the rest of the world, respectively. In terms of their potential to communicate during your PhD, they have some similarities which I will outline first, and some differences which I will treat later (see [here](#diffs)). Other platforms, such as [Slack](https://slack.com/) might be used in particular labs. These can be highly efficient ways of communicating with groups.

Just like [emails](#keepemails), these social messaging apps can be kept and archived. This means that you can use them as a means of documenting [due diligence](#deliverables). Make sure that you have your advisor's consent to do this. They, or your institutional policies, may insist that you inform them in an email.

It is important to bear in mind that these social media apps can have an adverse effect on your mental health [@montag2018multipurpose], and even become addictive [@hou2017excessive]. If you think that you need help because you are spending too much time checking your social media apps, then please get in touch with your university's psotgraduate student support.

### Direct messaging

It may be appropriate and acceptable to use a messaging app to communicate directly with your advisor. For example, if you have an [urgent request](#urgent). Many of the points discussed [above](#email) also apply here, and some are even more relevant:

 - Although a message may be received instantly, this does not mean that your advisor (or any other academic) may have the time to respond to you immediately. You need to give people time to respond. In particular this is especially important in evenings, weekends and holiday periods. 
 - As there is an expectation that direct messaging receives a quick answer, please try to avoid sending direct messages outside of office hours - unless this has been specifically sanctioned
 - If you are new at communicating with your advisor, ask them whether they prefer to use direct messaging over email. 
 - Having someone's direct messaging details means that you can call them and speak to them directly on their mobile device. Ask permission in the chat before you do this. If you are communicating with your advisor, ask them specifically about whether they are happy with this type of communication - don't just assume it's ok.
 - Rather like email, it is possible to [forward content](#fwd) from WeChat and WhatsApp to other uses. Be sure that this is appropriate and sanctioned before doing so. Alternatively, assume that everything that is shared with you is confidential unless you are given specific permission to share. 
 - Do remember to be polite and thankful when people help you through direct messages.

### Group messaging

In general, I would advise to learn the unwritten rules of the group through observation over time. Here are some issues that I have come across with student messaging groups.

Many students love to use gifs to display their emotions. I would suggest that this is inappropriate in most instances for two reasons: first they take up a lot of space and demand that anyone trying to catch up needs to scroll through a lot more content; second they are designed to be distracting and that is not the aim of communicating in these groups.

Keep your messages on topic. Because every message that you (or anyone else) sends on this group platform will send a notification to all subscribers, it is important that you keep any messages that you send relevant to everyone in the group. Avoid sending messages that are about social events (e.g. party invites) unless this has already been sanctioned and (importantly) everyone is invited.

#### In your postgraduate classes

Joining a group for your class may be mandatory in some institutions, or at the very least the only way in which you will learn about tasks and deadlines for postgraduate classes. Some postgraduate classes are experimenting with platforms that are only mediated through groups, such as Problem-Based Learning [@luo2021wechat]. In these instances you will have to use the app.

Before you post a question on the group, make sure that someone hasn't just already asked the same question. Use the search functions and see if there's anything that you have missed.

In your classes, it would be inappropriate to add someone who isn't in the class to the group. This might be seen as being nosey or snooping on class discussions.

If you are in doubt about what can and cannot be done in a class group, consult your lecturer or ask your class rep to consult them.

#### In your lab groups

Your lab group may also have a group messaging. Generally, these are extremely efficient ways to disseminate information to a group. Be aware of the general issues with these platforms and make sure that if you want to try to do something new on your lab group, have the blessing of your advisor or senior lab member before posting.

If your group message is intended for one person in particular, then rather don't bring it to the entire group. In the same way, if someone has what you believe to be inappropriate behaviour in the lab try to take this problem to the lab PI or bring it up in a lab meeting (as appropriate) rather than posting about it in the group. Calling out bad behaviour in a chat group is sometimes counter-productive and can be upsetting for many lab members.

### Desktop versions

Both WeChat and WhatsApp have desktop versions of their platforms, and these may replace the use of email in some places. For example, in China WeChat has almost entirely replaced email in many institutions such that sending an [email](#email) may elicit no response as the inbox is simply not checked. Note that this is not the case in Europe, North America and elsewhere. Institutions may well specifically forbid the official use of social media apps for official communications, so be aware of the policies at your institution.

### Turning off notifications

It is understandable that you (and others) in a group chat will want to turn off notifications if their mobile devices are pinging too often. If you do turn off your notifications for important class or lab groups, make sure that you look through the content at least once a day (excepting holidays).

### Differences {#diffs}
There are some important differences between WeChat and WhatsApp. At the time of writing, these differences are diminishing.

#### WeChat

WeChat has become a platform that offers far more than messaging and calls [@montag2018multipurpose]. You can do a lot more on it, as it effectively works as a browser. Thus within your WeChat postgraduate class group, there may be scope to include parts of the learning onto the app. All of this makes it more likely that WeChat will be used as a part of your learning experience. Indeed, there is even evidence that using WeChat responsibly for knowledge sharing at work promotes creativity [@zhang2021excessive].

Another important difference is that WeChat is often linked to someone's bank card and so can be used for transferring cash. Although this seems unlikely to be used in your PhD situation, always take care when doing this and find out before hand whether WeChat evidence for payments can be used for accounting purposes in your institution.

#### WhatsApp

At the moment, WhatsApp does not have the extended uses of the WeChat platform.

Having someone's WhatsApp contact details means that you will have their mobile phone number (or at least the one used for their account which may not be the SIM in their phone). Use this information with respect and don't share it with third parties unless you have consent.

## Other Social Media platforms

During your PhD studies you may make use of other social media platforms to communicate such as [Atlmetrics](https://howtopublishscience.org/altmetrics.html) [e.g., see @measey2022how]. As ever, use social media with caution and professionally. Respect the rights of you advisor and lab mates to have their own private social media, just as you may well have your own.